Post-selection CI example, assymetric, not unimodal at 0

### 2016-10-21 analysis/IPvsEM.Rmd
Revised	efron.fcr.Rmd.

```{r simulation_setup}
timed_sims = function(ash.args,nsim = 20,nmin = 100,nmax = 1000,betasd = 1) {
  n = 10^seq(log10(nmin),log10(nmax),length = nsim)
  elapsed.time = rep(0,nsim)
  for (i in 1:nsim) {
    set.seed(i)
    with(ash.args,cat(sprintf("%6s %13s %3d\n",method,optmethod,i)))
    betahat = rnorm(n[i],sd = sqrt(1 + betasd^2))
    elapsed.time[i] =
      system.time(do.call(ash,args = modifyList(ash.args,
        list(betahat = betahat, sebetahat = 1))))[3]
  }
  return(data.frame(elapsed.time = elapsed.time,seed = 1:nsim,n = n))
}
```

@@

```{r model_fitting}
df = data.frame()
cat("method --optmethod-- sim\n")
for(method in c("fdr","shrink")) {
  for(optmethod in c("mixIP","cxxMixSquarem")) {
    df = rbind(df,data.frame(method = method,optmethod = optmethod,
           timed_sims(list(method = method,optmethod = optmethod),
                           nsim = 50,nmin = 10,nmax = 100000)))
  }
}
cat("\n")

>>>>

```{r simulation_setup}
timed_sims = function(ash.args,nsim = 20,nmin = 100,nmax = 1e3,betasd = 1) {
  n = 10^seq(log10(nmin),log10(nmax),length = nsim)
  elapsed.time = rep(0,nsim)
  for (i in 1:nsim) {
    set.seed(i)
    with(ash.args,cat(sprintf("%6s %13s %3d\n",method,optmethod,i)))
    betahat = rnorm(n[i],sd = sqrt(1 + betasd^2))
    elapsed.time[i] =
      system.time(do.call(ash,args = modifyList(ash.args,
        list(betahat = betahat, sebetahat = 1))))[3]
  }
  return(data.frame(elapsed.time = elapsed.time,seed = 1:nsim,n = n))
}
```

@@

```{r model_fitting}
df = data.frame()
cat("method --optmethod-- sim\n")
for(method in c("fdr","shrink")) {
  for(optmethod in c("mixIP","cxxMixSquarem")) {
    df = rbind(df,data.frame(method = method,optmethod = optmethod,
           timed_sims(list(method = method,optmethod = optmethod),
                           nsim = 50,nmin = 10,nmax = 1e5)))
  }
}
cat("\n")

### 2016-10-21 analysis/ash_npmle.Rmd
Revised	efron.fcr.Rmd.

Now we fit the NPMLE using 'ashr' and compare with the REBayes solution.
```{r fit_ash_model}
grid = seq(from = min(z$x),to = max(z$x),length = 1000)
k    = length(grid)
y.ash.npmle =
  ash(y,1,g = unimix(pi=rep(1/(k-1),(k-1)),a = grid[-k],b = grid[-1]),
      method = "shrink")

>>>>

Now we fit the NPMLE using 'ashr' and compare with the REBayes solution.
```{r fit_ash_model}
grid = seq(from = min(z$x),to = max(z$x),length = 1000)
k    = length(grid)
y.ash.npmle =
  ash(y,1,g = unimix(pi = rep(1/(k-1),(k-1)),a = grid[-k],b = grid[-1]),
      method = "shrink")

### 2016-10-21 analysis/efron.fcr.Rmd
Revised	efron.fcr.Rmd.

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

@@

This example comes from Efron (2008) p16 when examining the false coverage rate (FCR). 
I selected this example because the distribution of the non-zero effect is
highly assymetric and not at all unimodal at zero, both issues a referee asked me
to elaborate on. Specifically, the distribution of the non-zero effects is N(-3,1).
Here I simulate data, and apply ash (with the "halfuniform" option to allow for asymmetric g).

```{r}
  set.seed(10)
  nsamp=10000
  altmean=-3
  mu0 = rep(0,nsamp)
  mu1 = rnorm(nsamp,altmean,1)
  comp = rbinom(nsamp,1,0.1)
  mu = ifelse(comp==0,mu0,mu1)
  z = rnorm(nsamp,mu,1)

  res.ash = ashr::ash(z,1,mixcompdist="halfuniform")

```

We can also run ash with the "true" g, to allow us to compare the lfsr, lfdr etc. 
```{r}
  true.g = ashr::normalmix(c(0.9,0.1),c(0,-3),c(0,1))
  res.ash.true = ashr::ash(z,1,g=true.g,fixg=TRUE)
```

Here we can see how the partition of $z$ scores compares with the truth. Note the effect of the unimodal assumption is to extend the inferred alternative distribution toward 0.
```{r}
  source("../R/nullalthist.R")
  par(mfcol=c(2,1))
  nullalthist(z,lfdr=res.ash.true$lfdr,main="true partition")
  nullalthist(z,lfdr=res.ash$lfdr,main="inferred partition")
```

Comparing the inferred Posterior Means, lfdr, and lfsr with the true values of these quantities, we find reassuringly good correspondence.
```{r}
  par(mfcol=c(1,1))
  plot(res.ash.true$PosteriorMean,res.ash$PosteriorMean,xlab="Truth", ylab="ash.hu",main="Posterior Mean (inferred vs truth)",xlim=c(-6,1),ylim=c(-6,1))
  abline(a=0,b=1,col=2,lwd=3)

    plot(res.ash.true$lfdr,res.ash$lfdr,xlab="Truth", ylab="ash.hu",main="lfdr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
  abline(a=0,b=1,col=2,lwd=3)

  plot(res.ash.true$lfsr,res.ash$lfsr,xlab="Truth", ylab="ash.hu",main="lfsr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
  abline(a=0,b=1,col=2,lwd=3)

```

We can also run qvalue and locfdr. We see that locfdr perhaps performs a bit better than ash for the decomposition here, but the estimated local fdrs are pretty similar. Here qvalue does less well because of the asymmetry which we didn't take account of.
```{r}
library(qvalue)
library(locfdr)
res.locfdr = locfdr(z,nulltype = 0)
res.qvalue = qvalue(p=pchisq(z^2,df=1,lower.tail = FALSE))  
#res.pos.qvalue = qvalue(p=pchisq(z[z>0]^2,df=1,lower.tail = FALSE))  
#res.neg.qvalue = qvalue(p=pchisq(z[z<0]^2,df=1,lower.tail = FALSE))  
#df1= data.frame(Truth = res.ash.true$lfdr,Estimate=res.locfdr$fdr,Method="locfdr")
#df2= data.frame(Truth = res.ash.true$lfdr,Estimate=res.ash$lfdr,Method="ash.hu")
#df3 =data.frame(Truth = res.ash.true$lfdr,Estimate=res.qvalue$lfdr,Method="qvalue")
#df4 =data.frame(Truth = res.ash.true$lfdr[z<0],Estimate=res.neg.qvalue$lfdr,Method="qvalue (-)")
#df5 =data.frame(Truth = res.ash.true$lfdr[z>0],Estimate=res.pos.qvalue$lfdr,Method="qvalue (+)")

#df = rbind(df1,df2,df3,df4,df5)
#ggplot2::qplot(data=df,x=Truth,y=Estimate,facets= .~Method)

par(mfrow=c(1,3))
plot(res.ash.true$lfdr,res.ash$lfdr,xlab="Truth (lfdr)", ylab="ash.hu",main="ash.hu",xlim=c(0,1),ylim=c(0,1))
abline(a=0,b=1,col=2,lwd=3)
plot(res.ash.true$lfdr,res.locfdr$fdr,xlab="Truth (lfdr)", ylab="Estimate",main="locfdr",xlim=c(0,1),ylim=c(0,1))
abline(a=0,b=1,col=2,lwd=3)
plot(res.ash.true$lfdr,res.qvalue$lfdr,xlab="Truth (lfdr)", ylab="Estimate",main="qvalue",xlim=c(0,1),ylim=c(0,1))
abline(a=0,b=1,col=2,lwd=3)

```

# Uniform tail curtails Credible Intervals

The following plot compares the (symmetric-tail) 95% CIs from ash (red) for the "significant" observations with Bayes rule (green), similar to Figure 8 from Efron. Note that the lower 97.5% point is pretty accurate, but the upper 97.5% point is curtailed - presumably due, at least in part, to the short tails of the uniform mixture.
```{r}
  CImatrix= ashr::ashci(res.ash,level=0.95)
  BayesComparePlot=function(CImatrix,altmean=-3,...){
    plot(z,mu,xlim=c(-8,0),...)
    points(z[CImatrix[,1]],CImatrix[,5],col="red")
    points(z[CImatrix[,1]],CImatrix[,4],col="red")

    fdr = 0.9*dnorm(z)/(0.9*dnorm(z)+0.1*dnorm(z,altmean,sqrt(2)))
    o=order(z)
    upper = ifelse(fdr[o]<0.025,(z[o]+altmean)/2+qnorm(0.975+fdr[o])/sqrt(2),0)
    lines(z[o],upper,col="green",lwd=2)
    lines(z[o],(z[o]+altmean)/2-qnorm(0.975)/sqrt(2),col="green",lwd=2)
    abline(v=max(z[fdr<0.05]))
  }
  BayesComparePlot(CImatrix,main="CIs for highly asymmetric and non-unimodal-at-zero data")
```

@@

```{r}
  res.ash.VB = ashr::ash(z,1,mixcompdist="halfuniform",optmethod="mixVBEM")
  CImatrix.VB= ashr::ashci(res.ash.VB,level=0.95)
```

Again, we can compare results with Bayes rule
```{r}
  BayesComparePlot(CImatrix.VB, main="CIs for highly asymmetric and non-unimodal-at-zero data \n Variational Version")
```

>>>>

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk_options, include=FALSE}
library(ashr)
library(qvalue)
library(locfdr)
source("chunk-options.R")
```

@@

This example comes from Efron (2008) p16 when examining the false coverage rate
(FCR). I selected this example because the distribution of the non-zero effects
is highly assymetric and not at all unimodal at zero, both issues a referee
asked me to elaborate on. Specifically, the distribution of the non-zero
effects is N(-3,1). Here I simulate data, and apply ash (with the "halfuniform"
option to allow for asymmetric g).

```{r fit_halfuniform_model}
set.seed(10)
nsamp   = 1e4
altmean = -3
mu0     = rep(0,nsamp)
mu1     = rnorm(nsamp,altmean,1)
comp    = rbinom(nsamp,1,0.1)
mu      = ifelse(comp == 0,mu0,mu1)
z       = rnorm(nsamp,mu,1)

# Fit the model.
res.ash = ash(z,1,mixcompdist = "halfuniform")
```

We can also run ash with the "true" g, to allow us to compare the lfsr, lfdr,
etc.

```{r fit_normalmix_model}
true.g       = normalmix(c(0.9,0.1),c(0,-3),c(0,1))
res.ash.true = ash(z,1,g = true.g,fixg = TRUE)
```

Here we can see how the partition of $z$ scores compares with the truth. Note
the effect of the unimodal assumption is to extend the inferred alternative
distribution toward 0. Here, `nullalthist` plots a histogram of z scores,
highlighting the alternative distribution of z scores that is implied by
the localfdr values lfdr.

```{r plot_zscore_hist}
source("../R/nullalthist.R")
par(mfcol = c(2,1),cex.main = 1,font.main = 1)
nullalthist(z,lfdr = get_lfdr(res.ash.true),main = "true partition (res.ash.true)")
nullalthist(z,lfdr = get_lfdr(res.ash),main = "inferred partition (res.ash)")
```

Comparing the inferred posterior means, lfdr, and lfsr with the true values of
these quantities, we find reassuringly good correspondence.

```{r scatterplots}
par(mfrow = c(1,3))
plot(get_pm(res.ash.true),get_pm(res.ash),xlab = "Truth",ylab = "ash.hu",
     main = "Posterior Mean (inferred vs truth)",pch = 20,cex = 0.6,
     xlim = c(-6,1),ylim = c(-6,1))
abline(a = 0,b = 1,col = "orangered",lwd = 2,lty = "dotted")

plot(get_lfdr(res.ash.true),get_lfdr(res.ash),xlab = "Truth", ylab = "ash.hu",
     main = "lfdr (inferred vs truth)",pch = 20,cex = 0.6,
     xlim = c(0,1),ylim = c(0,1))
abline(a = 0,b = 1,col = "orangered",lwd = 2,lty = "dotted")

plot(get_lfsr(res.ash.true),get_lfsr(res.ash),xlab = "Truth", ylab = "ash.hu",
     main = "lfsr (inferred vs truth)",pch = 20,cex = 0.6,
     xlim = c(0,1),ylim = c(0,1))
abline(a = 0,b = 1,col = "orangered",lwd = 2,lty = "dotted")
```

## Comparison with qvalue and locfdr

We can also run qvalue and locfdr. We see that locfdr perhaps performs a bit
better than ash for the decomposition here, but the estimated local fdrs are
pretty similar. Here qvalue does less well because of the asymmetry which it
didn't take account of.

```{r comparison_to_other_methods}
res.locfdr = locfdr(z,nulltype = 0)
res.qvalue = qvalue(p = pchisq(z^2,df = 1,lower.tail = FALSE))

par(mfrow = c(1,3))
plot(get_lfdr(res.ash.true),get_lfdr(res.ash),pch = 20,cex = 0.6,
     xlab = "Truth (lfdr)",ylab = "ash.hu",main = "ash.hu",
     xlim = c(0,1),ylim = c(0,1))
abline(a = 0,b = 1,col = "orangered",lty = "dotted",lwd = 2)

plot(get_lfdr(res.ash.true),res.locfdr$fdr,pch = 20,cex = 0.6,
     xlab = "Truth (lfdr)",ylab = "Estimate",main = "locfdr",
     xlim = c(0,1),ylim = c(0,1))
abline(a = 0,b = 1,lwd = 2,col = "orangered",lty = "dotted")

plot(get_lfdr(res.ash.true),res.qvalue$lfdr,pch = 20,cex = 0.6,
     xlab = "Truth (lfdr)",ylab = "Estimate",main = "qvalue",
     xlim = c(0,1),ylim = c(0,1))
abline(a = 0,b = 1,lwd = 2,col = "orangered",lty = "dotted")
```

## Uniform tail curtails Credible Intervals

The following plot compares the (symmetric-tail) 95% CIs from ash (red) for
the "significant" observations with Bayes rule (green), similar to Figure 8
from Efron. Note that the lower 97.5% point is pretty accurate, but the
upper 97.5% point is curtailed - presumably due, at least in part, to the
short tails of the uniform mixture.

```{r plot_credints}
CImatrix = ashci(res.ash,level = 0.95)
BayesComparePlot = function(CImatrix, altmean = -3,...) {
  plot(z,mu,xlim = c(-8,0),pch = 20,cex = 0.6,...)
  points(z,CImatrix[,1],col = "orangered",cex = 0.6,pch = 20)
  points(z,CImatrix[,2],col = "orangered",cex = 0.6,pch = 20)

  fdr   = 0.9*dnorm(z)/(0.9*dnorm(z) + 0.1*dnorm(z,altmean,sqrt(2)))
  o     = order(z)
  upper = ifelse(fdr[o] < 0.025,
                 (z[o] + altmean)/2 + qnorm(0.975 + fdr[o])/sqrt(2),0)
  lines(z[o],upper,col = "limegreen",lwd = 2)
  lines(z[o],(z[o] + altmean)/2 - qnorm(0.975)/sqrt(2),col = "limegreen",lwd = 2)
  abline(v = max(z[fdr < 0.05]))
}
par(mfrow = c(1,1))
BayesComparePlot(CImatrix,
  main = "CIs for highly asymmetric and non-unimodal-at-zero data")
```

@@

```{r vb_model_fitting}
res.ash.VB  = ash(z,1,mixcompdist = "halfuniform",optmethod = "mixVBEM")
CImatrix.VB = ashci(res.ash.VB,level = 0.95)
```

Again, we can compare results with Bayes rule:

```{r plot_vb_credints}
BayesComparePlot(CImatrix.VB,
  main = paste("CIs for highly asymmetric and non-unimodal-at-zero data\n",
              "Variational Version"))
```
